Testing and Isolation Efficacy; Insights from a Simple Epidemic Model

### 2021-01-05 SIR_manuscript.tex
updated intro

The early dynamics of an epidemic are determined by $\Rnum$. This index is defined as the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}. In the early stages of an epidemic the number of infected individuals is expected grow exponentially over time when $\Rnum>1$. To control the spread of disease in the population, $\Rnum$ is required to be below unity.

Here, we develop a model to incorporate epidemic processes and testing processes mechanistically to study the potential effect of testing strategies and test speed on epidemic control. We investigate the effect of underlying parameters on the basic reproduction number, $\Rnum$. While focused testing strategies are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is viral spread is expected to be stronger for a slow test than for a fast test. This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).

>>>>

The early dynamics of an epidemic are determined by $\Rnum$. This index is defined as the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}. In the early stages of an epidemic the number of infected individuals is expected grow exponentially over time when $\Rnum>1$. To control the spread of disease in the population, $\Rnum$ is required to be below unity. \ali{now may be say what we expect from testing and isolation to reduce R0 and the suprise that the reporting time increase R0. }

Here, we develop a compartmental model to incorporate epidemic processes and testing processes mechanistically to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, reporting time and isolation on epidemic dynamic. We investigate the effect of underlying parameters of the model on the basic reproduction number, $\Rnum$. While targeted testing strategies are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is viral spread is expected to be stronger for a slow test than for a fast test. This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).
